---
title: "PCA"
author: "Carmen Galaz-García"
date: "1/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(ggfortify)  # for PCA
library(janitor)
```

```{r}
food <- read.csv(here("Task2_PCA","usda_nutrients.csv")) %>% 
  clean_names() %>%
  drop_na()
```


```{r}
cheeses <- food %>% 
  filter(str_detect(short_descrip, "CHEESE,"),   # select cheeses
         str_detect(food_group, "Dairy") ) %>% 
  select(!(id:short_descrip)) %>%                # select columns
  select(!(common_name:scientific_name)) %>% 
  #select(descrip:vit_e_mg) %>% 
  select(descrip:fiber_g) %>% 
  mutate(descrip = str_sub(descrip, start=9)) %>% 
  separate(descrip, 
           into=c("name", "type"), 
           sep=",", 
           extra="merge") %>%
  filter(!str_detect(name,"fat") &
         !str_detect(name,"pasteurized") &
         !str_detect(name,"sodium")) %>% 
  filter(is.na(type) | str_detect(type,"fat", negate=TRUE)
           )
#  filter(!str_detect(name %in% c("fat","pasteurized","sodium"))

```


```{r}
cheeses <- food %>% 
  filter(str_detect(short_descrip, "CHEESE,"),   # select cheeses
         str_detect(food_group, "Dairy") ) %>% 
  select(!(id:short_descrip)) %>%                # select columns
  select(!(common_name:scientific_name)) %>% 
  select(descrip:fiber_g) %>% 
  mutate(descrip = str_sub(descrip, start=9)) %>% 
  separate(descrip, 
           into=c("name", "type"), 
           sep=",", 
           extra="merge") %>%
  filter(!str_detect(name,"fat|pasteurized|sodium")) %>% 
  filter(!str_detect(type,"fat|pasteurized|sodium|with|low") | is.na(type))
```


  filter(!str_detect(type,"fat"), 
         !str_detect(type,"pasteurized"),
         !str_detect(type,"sodium"),
         ) %>% 
  mutate(group = case_when(
    name == "mexican" ~ "mex",
    str_detect(type,"queso")~ "mex"
  )) %>% 
  relocate(group)
```

```{r}
cheeses_PCA <- cheeses %>% 
  select(3:8) %>%
  scale() %>% 
  prcomp

cheeses_PCA
```

```{r}
autoplot(cheeses_PCA,   # automatically recognizes PCA and does biplot
         data= cheeses,
         #colour = 'species', 
         loadings = TRUE,
         loadings.label=TRUE)  +
  theme_minimal()  # customize as any ggplot
```


